---
title: "Phloem Feeding Insects Inhibit Caterpillar Parasitism"
author: "Riley M. Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    code_folding: hide
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```





```{r Main_Code, include = F, cache = F}

# Load Packages
library(tidyverse) 
library(cowplot) 
library(lme4)
library(knitr)
library(forcats)
library(emmeans)
library(multcomp)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets:

para.by.spp <- read.csv("data/para.pf.spp.csv")

para.all.years <- read.csv("data/para.all.years.csv")

ant.data <- read.csv("data/pf.19.22.csv")

raw2022 <- read.csv("data/2022.raw.csv")

setup <- read.csv("data/Riley.sf.setup.2019.csv")

####################################################
#                    Functions                     #
####################################################

# parametric bootstrap confidence intervals for ant model:
bootfun.ant <- function(mod, dat){
  dat$newy <- simulate(mod)[[1]]
  mod_sim <- glm(newy ~ mem.treat * ant.treat, 
            data = dat, 
            family = binomial)
 plogis(c(coef(mod_sim)[1],
         coef(mod_sim)[2] + coef(mod_sim)[1],
         coef(mod_sim)[3] + coef(mod_sim)[1],
         coef(mod_sim)[4] + coef(mod_sim)[2]))
}

# Parametric bootstrap for the PF binomial model
bootfun <- function(mod, dat){
  dat$newy <- simulate( mod)[[1]]
  mod_sim <- glm(newy ~ treatment, 
            data = dat, 
            family = binomial, weights = yes + no)
 plogis(c(coef(mod_sim)[1], sum(coef(mod_sim))))
}


# Function to scale the range of values so that they are always positive or always negative, respectively

range01 <- function(x){
  (x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  }

```


# Overview

This analysis tests two hypotheses on the putative keystone effects of phloem-feeding insects. One keystone effect is the influence that phloem-feeders have on caterpillar-parasitoid interactions. These interactions may be mediated through ants (via symbiosis that attracts ants which can deter parasitoids) or plants (via phytohormonal crosstalk). The basic experimental design manipulates phloem-feeder presence by removing phloem-feeders or by removing and then replacing them. This experiment was carried out over 4 years, two of which were paired with an experimental ant manipulation (ants were excluded or ants had access). See *Anderson et al. _______* for details.

The statistical approach uses Fisher tests wherever possible for simplicity. However, we also employ the use of logistic regression models to analyze the factorial design in the ant experiment and to estimate predicted means and 95% confidence intervals. These statistics are estimated from parametric bootstraps of 1000 simulations from the posterior distributions of the binomial GLMs.


### Summary of Results
* Across four years, we experimentally manipulated `r nrow(para.all.years)` branches on white oak (*Quercus alba*) trees. 

* In 2019 and 2022 we implemented a factorial ant and phloem-feeder exclusion experiment. These experiments show that ants had no effect on parasitism. The distribution of parasitized caterpillars was essentially random across ant treatments (Fisher's exact test: *P* = 0.570, two-tailed test). Logistic regression of the two-way interaction confirms that the parasitism effect is independent of ant treatment.

* Across all four years, caterpillars were never parasitized in the presence of phloem-feeding insects (Fisher's exact test: *P* < 0.001, two tailed test).

```{r Data_Wrangling, echo = F, comment = ""}

# Experimental setup
setup1 <- setup %>% 
  select(branchID, type, treehoppers) %>% 
  distinct()

all.branches <- setup1 %>% 
  summarise(mean = mean(treehoppers),
            sd = sqrt(var(treehoppers))) %>% 
  mutate(Nbranches = nrow(setup1),
         NPFbranches = nrow(filter(setup1, treehoppers > 0)),
         freq.PFbranches = NPFbranches/Nbranches)

PF.branches <- setup1 %>% 
  filter(type == "experimental") %>% 
  summarise(mean = mean(treehoppers),
            sd = sqrt(var(treehoppers)))

cat.density <- setup %>% 
  select(branchID, counts) %>% 
  mutate(branchID = factor(branchID)) %>% 
  group_by(branchID) %>% 
  summarise(cats = sum(counts)) %>% 
  ungroup() %>% 
  summarise(mean = mean(cats),
            sd = sqrt(var(cats)))
  

# Ant data:
ant.data <- ant.data %>% 
  mutate(para = factor(para))

ant.19.22 <- ant.data %>% 
    select(ant.treat, para) %>% 
    group_by(ant.treat, para) %>% 
    tally()



ant.19.22.matrix <- data.frame(
  no = c(ant.19.22[[1,3]], ant.19.22[[3,3]]),
  yes = c(ant.19.22[[2,3]], ant.19.22[[4,3]]))

rownames(ant.19.22.matrix) <- c("Ant access", "Ant exclusion")

ant.tally <- ant.data %>% 
    group_by(mem.treat, ant.treat, para) %>% 
    tally()

filler <- data.frame(mem.treat = c("replaced", "replaced"),
                     ant.treat = c("access", "excluded"),
                     para = c("yes", "yes"),
                     n = c(0, 0)) %>% 
    mutate(mem.treat = factor(mem.treat),
           ant.treat = factor(ant.treat),
           para = factor(para))

pf.ant.fig.data <- bind_rows(ant.tally, filler)

pf.ant.fig.data <- pf.ant.fig.data %>% 
  mutate(mem.treat = factor(mem.treat),
         ant.treat = factor(ant.treat))

levels(pf.ant.fig.data$ant.treat) <- c("Ant access", "Ant exclusion") 
levels(pf.ant.fig.data$para) <- c("Unparasitized", "Parasitized")
levels(pf.ant.fig.data$mem.treat) <- c("removed", "replaced")


# Phloem-feeder parasitism data:

totals <- para.all.years %>% 
  mutate(treatment = factor(treatment)) %>% 
  group_by(treatment) %>% 
  summarise(no = sum(no),
            yes = sum(yes))


# Phloem-feeder parasitism data by caterpillar species:
para.pf1 <- para.by.spp %>% 
    select(cat.sp,
           total.rem,
           para.rem) %>% 
    mutate(yes = para.rem,
           no = total.rem - para.rem) %>% 
    select(cat.sp,
           yes,
           no) %>% 
    pivot_longer(!cat.sp,
                 names_to = "para",
                 values_to = "para_count") %>% 
    mutate(cat.sp = factor(cat.sp))

para.pf2 <- para.by.spp %>% 
    select(cat.sp,
           total.rep,
           para.rep) %>% 
    mutate(yes = para.rep,
           no = total.rep - para.rep,
           cat.sp = factor(cat.sp)) %>% 
    select(cat.sp,
           yes,
           no) %>% 
    pivot_longer(!cat.sp,
                 names_to = "para",
                 values_to = "para_count")


#########################################
# natural density of phloem-feeders

pf.dens <- raw2022 %>% 
  select(Site, Number.of.leaves, Phloem.feeders.found,
         Number.phloem.feeders, Density.phloem.feeders) %>% 
  mutate(membracid = if_else(grepl("mem", Phloem.feeders.found),
                             "yes", "no")) %>% 
  filter(membracid == "yes") %>% 
  mutate(mem.count = as.numeric(substr(Phloem.feeders.found,
                            start = 1, stop = 1))) %>% 
  mutate(leaf.area.m2 = 0.0100 * Number.of.leaves,
         mem.density = mem.count / leaf.area.m2)



total.mem.dens <- pf.dens %>% 
  summarise(mean = mean(mem.density),
            sd = sqrt(var(mem.density))) %>% 
  mutate(Site = "Overall")

col_order <- c("Site", "mean", "sd")

total.mem.dens <- total.mem.dens[, col_order]

site.mem.dens <- pf.dens %>% 
  group_by(Site) %>% 
  summarise(mean = mean(mem.density),
            sd = sqrt(var(mem.density)))

mem.density <- bind_rows(site.mem.dens, total.mem.dens)

```

# Natural membracid density in 2022 (membracids/m^2^)

```{r, membracid_density, echo = F}

kable(mem.density, digits = 3, format = "markdown")

```


# Natural membracid density in 2022 (membracids/branch)

```{r, membracid_density_per_branch, echo = F}

mems.branch <- pf.dens %>% 
  summarise(mean = mean(mem.count),
            sd = sqrt(var(mem.count)))

kable(mems.branch, digits = 3, format = "markdown")

```


# Setup metrics in 2019

## Membracid density (all sampled branches)
```{r mem_dens_all, echo = F}

kable(all.branches, digits = 3, format = "markdown")

```

## Membracid density (only branches with phloem-feeders)
```{r mem_dens_PF, echo = F}

kable(PF.branches, digits = 3, format = "markdown")

```

## Caterpillar density (all sampled branches)
```{r cat_dens_all, echo = F}

kable(cat.density, digits = 3, format = "markdown")

```


# Ants:

## 2019 & 2022 comparison of ant treatments on parasitism
```{r ant_pf_fisher_test, fig.width = 11, echo = F}

kable(ant.19.22.matrix, format = "markdown", digits = 2)

fisher.test(ant.19.22.matrix)

```

## Ant figure

```{r ant_pf_figure_manuscript, echo = F}

ggplot(pf.ant.fig.data,
       aes(x = mem.treat, y = n, fill = para)) +
    geom_col() +
    facet_wrap(~ ant.treat) +
    theme_classic(base_size = 20) +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
    scale_fill_manual(values = c("#56B4E9", "#D55E00")) +
    scale_y_continuous(limits = c(0, 45)) +
    labs(x = "Phloem-feeder treatment",
         y = "Caterpillar\ncount",
         fill = "Parasitism") +
    theme(legend.title.align = 0.5, strip.background=element_blank())


```

* Figure 1. Parasitism rate of caterpillars as a function of the factorial combinations of the phloem-feeder and ant treatments in Experiment 1.


## Logistic model of the factorial ant & pf treatments on parasitism
```{r ant_pf_logistic_model, echo = F}

ant.pf.mod <- glm(para ~ mem.treat * ant.treat,
                   family = binomial(),
                   data = ant.data)
ant.pf.mod

# p-value calculation is obstructed by data at the limits
# i.e. there are zero records of parasitism when phloem-feeders are replaced
# we apply two methods to estimate confidence intervals and p-values:
# 1st: bootstrap estimation from simulations of the posterior distribution
# 2nd: analysis of deviance (chosen because it is unaffected by the limits)

boot.ant <-  replicate(1000,
                       bootfun.ant(ant.pf.mod, dat=ant.data), 
                       simplify = TRUE)

plotres.ant <-  t(apply(boot.ant, 1,
                        quantile,
                        c(0.5, 0.025, 0.975)))

plotres.ant <- data.frame(
    "treatment" = c("PF- & Ant+",
                    "PF+ & Ant+",
                    "PF- & Ant-",
                    "PF+ & Ant-"), 
    plotres.ant)



names(plotres.ant)[-1] <-  c("parasitized", "lcl", "ucl")

rownames(plotres.ant) <- NULL

## get a p-value
# bootfun_rb <- function(mod){
#   simdat <- mod$data
#   simdat$para <- simulate(mod)[,1]
#   newmod <- update(mod, data =simdat)
#   return(coef(newmod))
# }


# boot_rb <- replicate(999, bootfun_rb(ant.pf.mod), simplify =TRUE)
# pvals <- 1 - (2*abs(apply(boot_rb, 1, function(x) mean(x>0)) - 0.5))
# round(pvals, 4)

# summary(ant.pf.mod) ## ant and intercept are similar to summary!
## checking the cis line up
 
# apply(boot_rb, 1,  quantile, c(0.025, 0.975)) ## note that antexcluded:memtrt is
## the reverse of the ant treatment - what we'd expect,

## Turns out the car package has an auto-magic way of getting bootstrapping
## glm objects. It estimates and accounts for bias etc. Trying that
# boot_car <- car::Boot(ant.pf.mod, R = 999 )
# car::Confint(boot_car, type = "bca")
## the package boot.pval automatically gives pvals if we want
#boot.pval::boot_summary(ant.pf.mod) ## to check against another approach

## Perhaps the most straightforward approach would be to do an analysis of 
## deviance, which isn't affected by the limits., 

AoD.results.ants <- car::Anova(ant.pf.mod)


```

* Analysis of deviance results: 
`r kable(AoD.results.ants, format = "markdown", digits = 4)`

## Ant & PF means and CIs figure

This figure is a visual representation of the logistic model approach (as opposed to Fisher's test). Here, we model the probability of parasitism as a function of ant treatment, phloem-feeder treatment, and their two-way interaction. Points and whiskers are means and 95% confidence intervals estimated from a parametric bootstrap from 1000 simulations of the posterior distribution of the binomial GLM. 
```{r ant_pf_logistic_model_parametric_CIs_fig, echo = F}

# model:
em1 <- emmeans(ant.pf.mod, ~mem.treat * ant.treat)


em1cld <- cld(em1, type = "response",
              letters = "abcde")

em1cld <- as_tibble(em1cld)

ggplot(plotres.ant, aes(x = treatment, y = parasitized,
                        ymin = lcl, ymax = ucl)) +
    geom_pointrange() +
  theme_classic(base_size = 22) +
  labs(x = "Phloem-feeder & Ant treatment",
       y = "Probability of Parasitism")

em1cld <- em1cld %>% 
  unite(x, mem.treat, ant.treat, sep = "_")

ggplot(em1cld, aes(x = x, y = prob)) +
  geom_point(size = 2) + 
  geom_errorbar(aes(ymin = prob - SE,
                    ymax = prob + SE),
                width = 0.2) +
  ylim(0, .75) +
  


```

* Caterpillars were only ever parasitized in the absence of phloem-feeders, and this effect was independent of ants. 



# Phloem-feeders:

## 2019-2022 parasitism analysis of phloem-feeders via Fisher's Exact Test

```{r para_fisher_test_19_20_21_22, echo = F}


kable(totals, format = "markdown", digits = 2)

fisher.test(totals[,2:3])


```


## 2019-2022 parasitism analysis of phloem-feeders via binomial GLM
```{r para_glm_19_20_21_22, echo = F}

# modeling
pf.mod <- glm(yes/(yes + no) ~ treatment, 
            data = para.all.years, 
            family = binomial, weights = yes + no)


boot.pf <-  replicate(1000,
                    bootfun(pf.mod, dat = para.all.years),
                    simplify = TRUE)

plotres <-  t(apply(boot.pf, 1, quantile, c(0.5, 0.025, 0.975)))

plotres <- data.frame("treatment" = c("Removed", "Replaced"), plotres)

names(plotres)[-1] <-  c("parasitized", "lcl", "ucl")

car::Anova(pf.mod)
head(para.all.years)
```

## Phloem-feeder figure

```{r para_fig_glm_CIs, fig.width = 5, echo = F}

ggplot(plotres, aes(x = treatment, 
                     y = parasitized, 
                     ymin = lcl,
                     ymax = ucl)) +
  geom_pointrange() +
    labs(x = "Phloem-feeder treatment",
         y = "Probabilty of parasitism") +
  theme_classic(base_size = 22) +
  scale_x_discrete(labels = c("removed", "replaced"))+
  expand_limits(y = .45)


```

* Figure 2. Fisher’s exact test showed that phloem-feeder replacement completely suppressed parasitism of caterpillars at five field sites over four years (N^total_caterpillars^ = 173, Experiments 1 and 2). Points and whiskers are means and 95% confidence intervals estimated from a parametric bootstrap of 1000 simulations from the posterior distribution of a binomial GLM.

## Parasitism by caterpillar species and PF treatment figure
```{r para_by_species_pf_treatment_fig, echo = F, fig.width = 10}

ggplot(para.pf1,
       aes(x = reorder(cat.sp, -para_count),
           y = para_count)) +
  geom_col(aes(fill = fct_rev(para)), width = 0.9) +
  geom_col(data = para.pf2,
           aes(y = -para_count,
               fill = para)) +
  coord_flip() +
  theme_cowplot() + 
  labs(x = "",
       y = "Caterpillar count",
       fill = "Parasitized") +
  geom_hline(yintercept = 0) +
  annotate(geom = "text",
           x = 21,
           y = -16, 
           label = "Phloem-feeder \nreplacement",
           color = "black",
           size = 7) +
  annotate(geom = "text",
           x = 21,
           y = 18, 
           label = "Phloem-feeder \nremoval",
           color = "black",
           size = 7) +
  theme(text = element_text(size = 22)) +
  scale_fill_manual(values = c("#D55E00", "#56B4E9")) +
  theme(legend.position = c(.8, .5))
    


```

* Figure 3. Distribution of caterpillar counts by species across phloem-feeder treatments (Experiments 1 and 2). Light grey bars indicate records of unparasitized caterpillars, whereas dark grey bars indicate parasitized caterpillars. Caterpillars were only parasitized when phloem-feeders were removed.


# Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


