---
title: "Phloem Feeding Insects Inhibit Caterpillar Parasitism"
author: "Riley M. Anderson & Robi Bagchi"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
  html_document:
    keep_md: yes
    theme: readable
  html_notebook:
    code_folding: hide
    theme: readable
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```





```{r libraries, include = F, cache = F}

# Load Packages
library(tidyverse) 
library(cowplot) 
library(knitr)
library(glmmTMB)
library(vegan)
library(parallel)
#library(tidymodels)
##library(lme4)
###library(forcats)
```

```{r load_data, cache = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

## Import datasets:
# experiment data:
dat1 <- read.csv("data/data.19.20.21.22.csv")

# recovery rate data (caterpillars stocked):
rec.data <- read.csv("data/recovery.csv")

# volatile data:
chem <- read.csv("data/Mike_Trees_herbivores2.csv")
```
####################################################
#                    Functions                     #
####################################################

```{r functions, cache = F}
# Function to scale the range of values so that they are always positive or always negative, respectively:

range01 <- function(x){
  (x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  }


# Function to check for overdispersion in poisson and binomial models:

overdispersion_test <- function(model, type = "pearson"){
    
    # Get the pearson residuals
    residuals <- resid(model, type = type)
    
    # Get the residual degrees of freedom of the model
    df <- df.residual(model)
    
    # Sum of residual deviance
    dev <- sum(residuals ^ 2)
    
    # Overdispersion = sum of squared residuals / residual degrees of freedom
    ratio <- round(dev / df, 3)
    
    # P-value 
    pvalue <- round(pchisq(dev, df, lower.tail = FALSE), 3)
    
    # Get the formula
    f = paste(as.character(formula(model))[2:3], collapse = " ~ ")
    
    # Get the model name
    name <- deparse(substitute(model))
    cat("Overdispersion ratio for model:", name, "\nformula:", f, 
        "\n\nAcceptable range: 1 - 1.4\nOverdispersion ratio:",
        ratio, " df:", df, " p =", pvalue, "\n", 
        ifelse(pvalue < 0.05, "Data are overdispersed\n", 
        "Data are not overdispersed\n"))
    
    # Return all the parameters
    return(c(ratio = ratio, deviance = dev, df = df, pvalue = pvalue))
    
}

# Function to generate ellipses dataset:

veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

###########
# Seed
set.seed(321)

```


# Overview


```{r Data_Wrangling, echo = F, comment = ""}
# experimental data:
dat1 <- dat1 %>% 
  mutate(across(c(branchID : para), as.factor),
         para = as.numeric(para) - 1,
         site = fct_recode(site, "PR" = "MS"))


# recovery rate data:
rec.data <- rec.data %>% 
  mutate(across(c(year:ant_treatment), as.factor),
         not_para = recovered - parasitized,
         prop_para = parasitized/recovered,
         prop_recovered = recovered/stocked)

# volatile data:
chem <- chem %>% 
  mutate(treatment.1 = ifelse(treatment.1 == "minus",
                              "removed", "replaced"))


```


## Branch level model for parasitism

```{r glmmtmb_branch, cache = TRUE, message=FALSE}
n_distinct(dat1$branchID); n_distinct(dat1$year); n_distinct(dat1$site);
n_distinct(dat1$cat_species) ## Enough replication for branch and cat_sp to be
# random. year and site need to be fixed.

branch_tmbmod <- glmmTMB(para ~ pf_treatment * ant_treatment +
                    year + site + (1|branchID) + (1|cat_species), 
                    family = binomial, data = dat1,
                    contrasts=list(year = "contr.sum", site = "contr.sum"))
summary(branch_tmbmod) ## Hauck-Donner effect due to complete separation.

## Using a parametric bootstrap to get around the problem
branch_tmbsims <- simulate(branch_tmbmod, 999)

## set up cluster for parallel computation
cl <- makeCluster(detectCores()-1)
clusterExport(cl, list("branch_tmbmod", "dat1"))
clusterEvalQ(cl, library(glmmTMB))
branch_tmbboot <- parSapplyLB(cl, branch_tmbsims, function(d){
  simmod <- refit(branch_tmbmod, newresp=d[,1])
  return(simmod)}, simplify=FALSE)

stopCluster(cl)

branch_tmbcis <- cbind(Est = fixef(branch_tmbmod)$cond,
                       mean_est = apply(sapply(branch_tmbboot, 
                                    function(m) fixef(m)$cond), 1, mean),
                       t(apply(sapply(branch_tmbboot, 
                                      function(m) fixef(m)$cond), 1,
                       quantile, c(0.025, 0.975))))
                       
kable(branch_tmbcis)

## analysis of deviance
## remove interaction
branch_tmbmod_main <- update(branch_tmbmod, ~.-ant_treatment:pf_treatment)
# test main effects (always against model with all main effects, so type 2 SS)
branch_tmbmod_pf <- update(branch_tmbmod_main, ~.-pf_treatment)
branch_tmbmod_ant <- glmmTMB(para ~ pf_treatment +
                               year + site +
                               (1|branchID) + (1|cat_species), 
                    family = binomial, data = dat1,
                    contrasts=list(site = "contr.sum"))
## update doesn't work with glmmTMB and multilevel factors
branch_tmbmod_y <- glmmTMB(para ~ pf_treatment + ant_treatment +
                    site + (1|branchID) + (1|cat_species), 
                    family = binomial, data = dat1,
                    contrasts=list(site = "contr.sum"))
branch_tmbmod_s <- glmmTMB(para ~ pf_treatment + ant_treatment +
                             year +  (1|branchID) + (1|cat_species), 
                           family = binomial, data = dat1,
                            contrasts=list(year = "contr.sum"),
                           se = FALSE) 
### this model doesn't converge due to Hessian computation issues. But for a 
### andev we don't need the SEs, so not computing here to allow some convergence

AIC(branch_tmbmod, branch_tmbmod_main, 
                    branch_tmbmod_pf, branch_tmbmod_ant,
                    branch_tmbmod_y, branch_tmbmod_s)

## dropping pf clearly increases AIC (by 19 units from main effect model).
## make analysis of deviance table (a bit tedious)
andev <- function(m0, m1){
  tab <- as.data.frame(anova(m0, m1))
  tab <- cbind(t(apply(tab[,2:5], 2, diff)), tab[2, 6:8] )
  tab[,4] <- tab[,4]*(-1)
  return(tab)}

andev_tab <- rbind(
  "Site" = andev(branch_tmbmod_s, branch_tmbmod_main),
  "Year" = andev(branch_tmbmod_y, branch_tmbmod_main),
  "Phloem-feeders" = andev(branch_tmbmod_pf, branch_tmbmod_main),
  "Ants" = andev(branch_tmbmod_ant, branch_tmbmod_main),
  "Phloem-feeders x Ants" = andev(branch_tmbmod, branch_tmbmod_main))

kable(andev_tab, format = "markdown", digits = 3)
  
```


## Figure 2

### proportion recovered-all years
```{r Figure_2_prop, echo = F}
n_cats_treatment <- dat1 %>% 
  group_by(pf_treatment, ant_treatment) %>% 
  tally() %>% 
  rename(ncats = n)

fig2data <- dat1 %>% 
  mutate(para = factor(para)) %>% 
  group_by(pf_treatment, ant_treatment, para) %>% 
  tally() %>% 
  left_join(., n_cats_treatment,
            by = c("pf_treatment", "ant_treatment")) %>% 
  mutate(ant_treatment = case_when(
    ant_treatment == "access" ~ "Ant access",
    TRUE ~ "Ant exclusion"
  ))

fig2labels <- fig2data %>% 
  filter(para == "1") %>% 
  mutate(prop = round(100 * (n / ncats), 1),
         label = paste0(prop, "%", sep = ""))

fig2_prop <- fig2data %>%
  ggplot(aes(x = pf_treatment, y = n/ncats, fill = para)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ ant_treatment) +
  geom_text(data = fig2labels,
            label = fig2labels$label,
            nudge_y = -0.08) +
  labs(x = "Phloem-feeder treatment",
       y = "Prop. of caterpillars\nrecovered",
       fill = "") +
  scale_fill_manual(values = c("#56B4E9", "#D55E00"),
                    labels = c("unparasitized", "parasitized")) +
  theme_cowplot() +
  theme(legend.title.align = 0.5,
        strip.background = element_blank()) 

fig2_prop

```

## Recovery rates of stocked caterpillars

```{r recovery, echo = F}

rec.mod <- glm(prop_recovered ~ pf_treatment * ant_treatment +
                 site + year,
               family = quasibinomial(), # for overdispersion
               data = filter(rec.data, prop_recovered <= 1),
               # remove 1 record where recovery > stocked
               weights = stocked,
               contrasts = list(year = "contr.sum",
                                site = "contr.sum"))



kable(tidy(rec.mod), format = "markdown", digits = 3)

```

## Ants and Phloem-feeders:

### counts all years

```{r Figure_2_allyears, echo = F}

dat1 %>% 
  mutate(ant_treatment = case_when(
    ant_treatment == "access" ~ "Ant access",
    TRUE ~ "Ant exclusion"
  ),
  para = factor(para)) %>% 
  group_by(pf_treatment, ant_treatment, para) %>% 
  tally() %>% 
  ggplot(aes(x = pf_treatment, y = n, fill = para)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ ant_treatment) +
  scale_fill_manual(values = c("#56B4E9", "#D55E00"),
                    labels = c("unparasitized",
                               "parasitized")) +
  theme_cowplot(font_size = 14) +
  theme(legend.title.align = 0.5,
        strip.background = element_blank(),
        legend) +
  labs(x = "Phloem feeder treatment",
       y = "Caterpillar count",
       fill = "") +
  geom_text(data = fig2labels,
            label = fig2labels$label,
            nudge_y = -1.5)
```


### all years no ant effects

```{r Figure_2_no_ants, echo = F}

dat1 %>% 
  mutate(para = factor(para)) %>% 
  group_by(pf_treatment, para) %>% 
  tally() %>% 
  ggplot(aes(x = pf_treatment, y = n, fill = para)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#56B4E9", "#D55E00"),
                    labels = c("unparasitized",
                               "parasitized")) +
  theme_cowplot() +
  labs(x = "Phloem feeder treatment",
       y = "Caterpillar count",
       fill = "")

```



```{r new_pf_figure2, echo = F, fig.width = 10}

Atest <- dat1 %>% 
  mutate(para = factor(para)) %>% 
  ggplot(aes(x = year, y = para,
             color = pf_treatment, shape = ant_treatment)) +
  geom_jitter(height = .4, width = .22, size = 3) +
  scale_color_manual(values = c("#D57E00", "#56B4E9")) +
  scale_y_discrete(labels = c("not parasitized", "parasitized")) +
  labs(x = "Year", y = "",
       color = "Phloem-feeders",
       shape = "Ants") +
  geom_hline(yintercept = 1.5, linetype = "dashed") +
  theme_cowplot(font_size = 18) +
  coord_flip()

A <- dat1 %>% 
  mutate(para = factor(para)) %>% 
  ggplot(aes(x = year, y = para,
             color = pf_treatment, shape = ant_treatment)) +
  geom_jitter(height = .4, width = .15, size = 3) +
  scale_color_manual(values = c("#D57E00", "#56B4E9")) +
  scale_y_discrete(labels = c("", ""),
                   expand = c(.15, 0.3)) +
  labs(x = "Year", y = "",
       color = "Phloem-feeders",
       shape = "Ants") +
  geom_hline(yintercept = 1.5, linetype = "dashed") +
  theme_cowplot(font_size = 18) +
  coord_flip() +
  theme(legend.position = "none")

  
B <- dat1 %>% 
  mutate(para = factor(para)) %>% 
  ggplot(aes(x = site, y = para,
             color = pf_treatment, shape = ant_treatment)) +
  geom_jitter(height = .4, width = .15, size = 3) +
  scale_color_manual(values = c("#D57E00", "#56B4E9")) +
  scale_y_discrete(labels = c("not parasitized", "parasitized"),
                   expand = c(0.15, .3)) +
  scale_x_discrete(labels = c(" CK ", " CP ", " GH ",  " PR ")) +
  labs(x = "Site", y = "",
       color = "Phloem-feeders",
       shape = "Ants") +
  geom_hline(yintercept = 1.5, linetype = "dashed") +
  theme_cowplot(font_size = 18) +
  coord_flip() +
  theme(legend.position = "none")

grobs <- ggplotGrob(Atest)$grobs

legend <- grobs[[which(sapply(grobs, function(x) x$name) == "guide-box")]]

p1 <- cowplot::plot_grid(A, NULL, B, labels = c("A", "", "B"),
          ncol = 1, nrow = 3, rel_heights = c(1, -0.2, 1),
          greedy = T)

cowplot::plot_grid(p1, legend, ncol = 2,
          rel_widths = c(.45, 0.1),
          rel_heights = c(4, 4))

```

**Figure 2.** Caterpillar parasitism patterns across years (A) and sites (B) by experimental phloem-feeder manipulation (removal in orange, replacement in blue) and experimental ant treatment (ant access in circles, ant exclusion in triangles). Each point is a single caterpillar. Points are jittered for clarity. For detailed description of sites, tree replicates, and caterpillar sample sizes, see table S1.

## Supplemental table
```{r supp_table, echo = F}
rec.year <- rec.data %>% 
  select(-not_para, -prop_para, -prop_recovered) %>% 
  mutate(pfant = factor(paste(pf_treatment, ant_treatment, sep = "_"))) %>% 
  select(-pf_treatment, -ant_treatment, -site) %>% 
  group_by(year, pfant) %>%
  summarise(stocked = sum(stocked),
            recovered = sum(recovered),
            parasitized = sum(parasitized)) %>% 
  pivot_wider(names_from = pfant,
              values_from = c(stocked, recovered, parasitized))



rec.site <- rec.data %>% 
  select(-not_para, -prop_para, -prop_recovered) %>% 
  mutate(pfant = factor(paste(pf_treatment, ant_treatment, sep = "_"))) %>% 
  select(-pf_treatment, -ant_treatment, -year) %>% 
  group_by(site, pfant) %>%
  summarise(stocked = sum(stocked),
            recovered = sum(recovered),
            parasitized = sum(parasitized)) %>% 
  pivot_wider(names_from = pfant,
              values_from = c(stocked, recovered, parasitized))

rec.year.site <- bind_rows(rec.year, rec.site)

write.csv(rec.year.site,
          file = "data/recover.year.site.csv",
          row.names = F)

```



## Volatiles

```{r volatile_prep, echo = F, include = FALSE}

# NMDS
nmds <- metaMDS(chem[, 5:37], distance = "bray", k = 2)

# Add the "treatment" variable to the NMDS result
nmds_data <- data.frame(nmds$points, treatment = chem$treatment, 
                        exclusion = chem$treatment.1, 
                        taxon = chem$ploem.feeder,
                        sites = 1:16,
                        row.names = "sites")

# Calculate centroid for each treatment
centroids <- nmds_data %>% 
  group_by(exclusion, taxon, treatment) %>%  
  summarize(MDS1 = mean(MDS1), MDS2 = mean(MDS2))

# ellipses
nmds_data <- nmds_data %>% 
  unite(treatment, c("exclusion", "taxon"))

nmdsmeta <- data.frame(
  MDS1 = nmds$points[,1],
  MDS2 = nmds$points[,2],
  group = nmds_data$treatment
)

nmdsmeta <- nmdsmeta %>% 
  mutate(group = factor(group))

plot(nmds)
ord <- ordiellipse(nmds,
                   nmds_data$treatment,
                   display = "sites",
                   kind = "se", conf = 0.95, label = T)

df_ell <- data.frame()

for(g in levels(nmdsmeta$group)){
  df_ell <- rbind(df_ell,
                  cbind(
                    as.data.frame(with(
                      nmdsmeta[nmdsmeta$group==g,],
                      veganCovEllipse(ord[[g]]$cov,
                                      ord[[g]]$center,
                                      ord[[g]]$scale)))
                                ,group=g))
}

df_ell <- df_ell %>% 
  mutate(exclusion = factor(if_else(grepl("replaced", group),
                             "replaced", "removed")),
         taxon = factor(if_else(grepl("membracid", group),
                         "membracid", "coccid")))

nmds_data <- nmds_data %>% 
  separate_wider_delim(treatment, delim = "_",
                       names = c("exclusion", "taxon"))

cent.labels <- centroids %>% 
  ungroup() %>% 
  mutate(letter = c("a", "a", "ab", "b"))

```

### Volatile Plot
```{r Volatile_plot, echo = F}

ggplot(nmds_data, aes(x = MDS1, y = MDS2, 
                      color = taxon,
                      shape = exclusion)) +
  geom_point(data = nmds_data, size = 1, show.legend = F) +
  geom_point(data = centroids, aes(shape = exclusion),
             size = 4) +
  geom_polygon(data = df_ell,
               aes(x = NMDS1, y = NMDS2,
                   color = taxon, linetype = exclusion), 
               alpha = 0.3, fill = NA,
               show.legend = F) +
  labs(x = "NMDS1", y = "NMDS2",
       shape = "Treatment",
       color = "Taxon") +
  scale_shape_manual(values=c(1, 16)) +
  theme_cowplot(font_size = 16) +
  scale_color_manual(values = c("#D55E00", "#56B4E9")) +
  #guides(shape = guide_legend(override.aes = list(size = 3))) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  geom_text(data = cent.labels,
            label = cent.labels$letter,
            nudge_y = 0.06, nudge_x = -0.06,
            color = 'black')

```

```{r volatile_inference}
chem$block <- rep(1:8, each = 2)
chem <- rename(chem, "pf_treatment" = "treatment.1",
               "pf_species" = "ploem.feeder")
names(chem)
vmod <- adonis2(chem[, 5:37] ~ pf_treatment * pf_species, 
                strata = chem$block, 
                data = chem, permutations=999)

vmod_m <- adonis2(chem[, 5:37] ~ pf_treatment + pf_species, 
                strata = chem$block, 
                data = chem, permutations=999)


print(vmod) ## both  pf_treatment and pf_species now have significant
## effects on voc composition.

library(randomForest)
rfmod_pf <- randomForest(y = as.factor(chem$pf_treatment), 
                      x = select(chem, starts_with("C")), 
                      strata = chem$block, importance = TRUE, nPerm = 999)
varImpPlot(rfmod_pf)
imp_vals_pf <- data.frame(importance(rfmod_pf))
imp_vals_pf <- imp_vals_pf |> arrange(desc(MeanDecreaseGini))
## find values with within the top 10% of importance values
imp_vals_toppf <- filter(imp_vals_pf,
                         MeanDecreaseGini > quantile(MeanDecreaseGini, 0.9)) 
imp_vals_toppf

## differences between taxa
rfmod_sp <- randomForest(y = as.factor(chem$pf_species), 
                      x = select(chem, starts_with("C")), 
                      strata = chem$block, importance = TRUE, nPerm = 500)
varImpPlot(rfmod_sp)
imp_vals_sp <- data.frame(importance(rfmod_sp))
imp_vals_sp <- imp_vals_sp |> arrange(desc(MeanDecreaseGini))
## find values with within the top 10% of importance values
imp_vals_topsp <- filter(imp_vals_sp,
                         MeanDecreaseGini > quantile(MeanDecreaseGini, 0.9)) 
pf_sp_comp <- cbind(imp_vals_sp, imp_vals_pf[rownames(imp_vals_sp),])
names(pf_sp_comp)[c(4, 8)] <- c("gini_sp", "gini_pf")
cor(pf_sp_comp[,c("gini_sp", "gini_pf")]) ## low correlation.

chem_imp <- select(chem, pf_treatment, pf_species, block, 
                  rownames(imp_vals_toppf), rownames(imp_vals_topsp))

pf_aovs <- sapply(rownames(imp_vals_toppf), function(voc){
  # form <- paste(voc, "~  pf_treatment * pf_species + (1|block)")
  # mod <- lmerTest::lmer(as.formula(form), data = chem_imp)
  form <- paste(voc, "~  pf_treatment * pf_species + Error(block)") 
  mod <- aov(as.formula(form), data= chem_imp)
}, simplify = FALSE)


sp_aovs <- sapply(rownames(imp_vals_topsp), function(voc){
  #form <- paste(voc, "~  pf_treatment * pf_species + (1|block)") 
  #mod <- lmerTest::lmer(as.formula(form), data = chem_imp)
  form <- paste(voc, "~  pf_treatment * pf_species + Error(block)") 
  mod <- aov(as.formula(form), data= chem_imp)
}, simplify = FALSE)

lapply(pf_aovs, function(m) kable((summary(m)[[2]][[1]])))
lapply(sp_aovs, function(m) kable(summary(m)[[2]][[1]]))

## if using lme4 version, but aov is simpler.
## singularities due to 0 variance among blocks.
## Not  a big issue 
# sapply(pf_aovs, function(m) VarCorr(m)$block)
# sapply(sp_aovs, function(m) VarCorr(m)$block)
# 
# sapply(pf_aovs, function(m) kable(anova(m, type =2)))
# sapply(sp_aovs, function(m) kable(anova(m, type =2)))
```




## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


